---
title: "Appendix"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

Import libraries 
```{r eval=FALSE}
library(forecast)
library(astsa)
library(ggplot2)
```

Import data. Have the csv in the same working directory as the code. 
```{r eval=FALSE}
stocks <- read.csv('stocks.csv') # Read in the data
stocks$Date <- as.Date(stocks$Date) # Change Dates to be date types
```

# Exploratory Data Analysis

Plotting data [Figure 1]
```{r eval=FALSE}
title1 <- 'Mediocre Social Network Apps Inc. Stock Price over Time'
figure1 <- ggplot(stocks) + geom_line(aes(x=Date, y=Price)) 
figure1 <- figure1 + labs(x = 'Time', y = 'Price ($)')  
figure1 <- figure1 + ggtitle(title1) 
figure1 <- figure1 + theme_bw() 
figure1
```

Fitting quadratic curve and find its residuals 
```{r eval=FALSE}
prices <- stocks$Price
time <- stocks$X
time_2 <- time ** 2
mod <- lm(prices~time + time_2)
stocks$resid <- mod$residuals
stocks$fitted <- mod$fitted.values
```

Plot the quadratic curve [Figure 2 (a)]
```{r eval=FALSE}
title2_a <- 'Quadratic Fit on Stock Price'
figure2_a <- ggplot(stocks) + geom_line(aes(x=Date, y=Price)) 
figure2_a <- figure2_a + geom_line(aes(x=Date, y=fitted), color = 'red') 
figure2_a <- figure2_a + labs(x = 'Time', y = 'Price ($)') 
figure2_a <- figure2_a + ggtitle(title2_a) 
figure2_a <- figure2_a + theme_bw()
figure2_a
```

Plot the residuals of the quadratic curve [Figure 2 (b)]
```{r eval=FALSE}
title2_b <- 'Residuals of Quadratic Model'
figure2_b <- ggplot(stocks) + geom_line(aes(x=Date, y=resid)) 
figure2_b <- figure2_b + labs(x = 'Time', y = 'Residual') 
figure2_b <- figure2_b + ggtitle(title2_b) 
figure2_b <- figure2_b + theme_bw()
figure2_b
```

# 2 Modeling a Deterministic Function of Time

Take the log of the data, fit the quadratic, and find the residuals
```{r eval=FALSE}
stocks$transform <- log(stocks$Price)
mod <- lm(stocks$transform~time + time_2)
stocks$transform_resid <-  mod$residuals
stocks$transform_fitted <- mod$fitted.values
```

Plot the quadratic fit on $X_t$ [Figure 3 (a)]
```{r eval=FALSE}
title3_a <- 'Quadratic fit on'~X[t]
figure3_a <- ggplot(stocks) + geom_line(aes(x=Date, y=transform)) 
figure3_a <- figure3_a + geom_line(aes(x=Date, y=transform_fitted), color = 'red') 
figure3_a <- figure3_a + labs(x = 'Time', y = 'Price (S)') + ggtitle(title3_a) + theme_bw()
figure3_a
```

Plot the residuals for the quadratic fit on $X_t$ [Figure 3 (b)]
```{r eval=FALSE}
title3_b <- 'Residuals of Quadratic Model on'~X[t]
figure3_b <- ggplot(stocks) + geom_line(aes(x=Date, y=transform_resid)) 
figure3_b <- figure3_b + labs(x = 'Time', y = 'Residual') + ggtitle(title3_b) + theme_bw()
figure3_b
```

Find and plot the first difference of $X_t$
```{r eval=FALSE}
stocks$diff_poly <- c(0, diff(stocks$transform_resid))
```

Plot the first difference of $X_t$ [Figure 4]
```{r eval=FALSE}
title4 <- '$\\nabla(\\epsilon_{t})$'
figure4 <- ggplot(stocks) + geom_line(aes(x=Date, y=diff_poly)) 
figure4 <- figure4 + labs(x = 'Time', y = 'Difference') + ggtitle(title4) + theme_bw()
figure4
```

Find $\nabla^2 X_t$ 
```{r eval=FALSE}
stocks$diff_2_log <- c(0, 0, diff(stocks$transform, differences = 2))
```

Plot $\nabla^2 X_t$ [Figure 5]
```{r eval=FALSE}
title5 <- '$\\nabla^{2}(X_{t})$'
figure5 <- ggplot(stocks) + geom_line(aes(x=Date, y=diff_2_log)) 
figure5 <- figure5 + labs(x = 'Time', y = 'Difference') + ggtitle(title5) 
figure5 <- figure5 + theme_bw()
figure5
```

Find $\nabla X_t$
```{r eval=FALSE}
stocks$diff_log <- c(0,diff(stocks$transform))
```

Plot of the first difference of the log data [Figure 6]
```{r eval=FALSE}
title6 <- '$\\nabla(X_{t})$'
figure6 <- ggplot(stocks) + geom_line(aes(x=Date, y=diff_log)) 
figure6 <- figure6 + labs(x = 'Time', y = 'Difference') + ggtitle(title6) 
figure6 <- figure6 + theme_bw()
figure6
```

# ARIMA Model Selection

Plot the ACF of $\nabla(\epsilon)$ [Figure 7 (a)]
```{r eval=FALSE}
title7_a <- 'ACF of $\\nabla$($\\epsilon_{t}$)'
acf(stocks$diff_poly, main = title7_a)
```

Plot the PACF of $\nabla(\epsilon)$ [Figure 7 (b)]
```{r eval=FALSE}
title7_b <- 'PACF of $\\nabla$($\\epsilon_{t}$)'
pacf(stocks$diff_poly, main = title7_b)
```

Plot the ACF of $\nabla(\epsilon)$ [Figure 8 (a)]
```{r eval=FALSE}
title8_a <- 'ACF of $\\nabla^{2}(X_{t})$'
acf(stocks$diff_2_log, main=title8_a)
```

Plot the PACF of $\nabla(\epsilon)$ [Figure 8 (b)]
```{r eval=FALSE}
title8_b <- 'PACF of $\\nabla^{2}(X_{t})$'
pacf(stocks$diff_2_log, main=title8_b)
```

The diagonistics of ARIMA(0,2,1) for the log data [Figure 9]
```{r eval=FALSE}
sarima(stocks$transform, 0, 2, 1)
```

The ACF and PACF of $\nabla X_t$ [Figure 10]
```{r eval=FALSE}
title10_a <- "ACF of $\\nabla(X_{t})"
acf(stocks$diff_log, main=title10_a)
```

The ACF and PACF of $\nabla X_t$ [Figure 10]
```{r eval=FALSE}
title10_b <- "PACF of $\\nabla(X_{t})"
pacf(stocks$diff_log, main=title10_b)
```

## Cross Validation Model Comparison

A function like sarmia.for but doesn't generate plots. Credit: https://stackoverflow.com/questions/55871256/can-i-prevent-sarima-for-from-plotting

```{r eval=FALSE}
sarima.noplot = function(x, ...) {
  png(tf<-tempfile())
  out <- sarima.for(x, ...)
  dev.off()
  file.remove(tf)
  return(out)
}
```

Calculates the MSE with cross-validation
```{r eval=FALSE}
mse <- function(p, d, q, P, D, Q, S) {
  start_year <- 254 
  end_year <- length(stocks$Price)
  sum_squared_errors = 0
  for (year in seq(start_year, end_year, 10)) {
    train_set <- window(stocks$Price, end=year-1)
    test_set <- window(stocks$Price, start=year, end=year + 9)
    forecast <- exp(sarima.noplot(log(train_set), 
                                 n.ahead=10, p = p, d = d, q = q, P = P, D = D, Q = Q, S = S)$pred)
    sum_squared_errors <- sum_squared_errors + sum((forecast - test_set)^2)
  }
  mse <- sum_squared_errors / (10*length(seq(start_year, end_year, 10)))
  return(mse)
}
```

Log Stabilization Transformed + ARIMA(0, 2, 1)
```{r eval=FALSE}
sqrt(mse(0,2,1,0,0,0,0))
```

Log Stabilization Transformed + ARIMA(0, 1, 0)
```{r eval=FALSE}
sqrt(mse(0,1,0,0,0,0,0))
```

Log Stabilization Transformed + Quadratic Trend + ARIMA(0, 1, 0)
```{r eval=FALSE}
start_year <- 254
end_year <- length(stocks$Price)
sum_squared_errors <- 0
for (year in seq(start_year, end_year, 10)) {
    train_set <- window(stocks$Price, end=year-1)
    test_set <- window(stocks$Price, start=year, end=year + 9)
    
    model4 <- lm(log(train_set) ~ poly(1:length(train_set), degree=2, raw=TRUE))
    test_matrix <- model.matrix( ~ poly((length(train_set) + 1):(length(train_set) + 10), 
                                        degree=2, raw=TRUE))
    trendforecast4 <- test_matrix %*% model4$coefficients
    temp <- sarima.noplot(model4$residuals, 
                          n.ahead = 10, p = 1, d = 0, q = 0, P = 0, D = 0, Q = 0, S = 0)$pred
    forecast4 <- exp(trendforecast4 + temp)
    sum_squared_errors <- sum_squared_errors + sum((forecast4 - test_set)^2)
}

sqrt(sum_squared_errors / (10*length(seq(start_year, end_year, 10))))
```

Information for ARIMA(0,1,0)
```{r eval=FALSE}
sarima(stocks$transform, 0, 1, 0)
```

Create predictions
```{r eval=FALSE}
last <- tail(stocks, 100)
forecast <- sarima.for(stocks$transform, n.ahead = 10, 0, 1, 0)
predictions <- exp(forecast$pred[1:10])
lower_bound_1 <- exp(forecast$pred[1:10] - forecast$se[1:10])
upper_bound_1 <- exp(forecast$pred[1:10] + forecast$se[1:10])
lower_bound_2 <- exp(forecast$pred[1:10] - 1.96*forecast$se[1:10])
upper_bound_2 <- exp(forecast$pred[1:10] + 1.96*forecast$se[1:10])
last <- last[c('Date', 'Price')]
last$Key <- 'Observations'
new_dates <- c("2019-10-01", "2019-10-02", "2019-10-03", "2019-10-04", 
               "2019-10-07", "2019-10-08", "2019-10-09", "2019-10-10", 
               "2019-10-11", "2019-10-14")
preds <- data.frame(Date = new_dates, Price = predictions, Key = 'Predictions')
upper_2 <- data.frame(Date = new_dates, Price = upper_bound_2,  Key = 'Upper_2')
lower_2 <- data.frame(Date = new_dates, Price = lower_bound_2,  Key = 'Lower_2')
upper_1 <- data.frame(Date = new_dates, Price = upper_bound_1,  Key = 'Upper_1')
lower_1 <- data.frame(Date = new_dates, Price = lower_bound_1,  Key = 'Lower_1')
df <- rbind(last, preds, upper_2, lower_2, upper_1, lower_1)
```

Plot prediction [Figure 11]
```{r eval=FALSE}
figure11 <- ggplot(df) + geom_line(aes(x=Date, y=Price, color = Key)) 
figure11 <- figure11 + labs(x = 'Time', y = 'Difference') + ggtitle('Predictions') 
figure11 <- figure11 + theme_bw() + theme(legend.position = 'none') 
figure11 <- figure11 + scale_colour_manual(values=c(
Observations='black', Predictions='red', Upper_1='deepskyblue', Lower_1='deepskyblue', 
Upper_2='deepskyblue3', Lower_2='deepskyblue3'))
figure11
```



